Spitzer, Gaia and the Potential of the Milky Way

### 2013-07-31 text/paper1/paper.tex
add bootstrap figure placeholder
kvj suggestions

\subsection{Gaia and the age of astrometry}
\label{sec:gaia}
The Gaia satellite \citep{gaia01}, planned for launch in late 2013, is
an astrometric mission which aims to measure the positions of billions
of stars with 10-100~$\mu$as accuracies. With over 50 observations of
the entire sky during the five-year lifetime of the mission, Gaia will
revolutionize the field of Galactic Astronomy by providing sufficient
proper motion accuracy to create full six-dimensional phase-space maps
of the Galaxy with $<$10\% distance errors for heliocentric distances
up to $\sim$6~kpc for RR Lyrae stars.

Figure~\ref{fig:gaia_errors} shows the Gaia end-of-mission distance
and tangential velocity error estimates for RR Lyrae. The velocity
uncertainty depends on the magnitude of the source, which depends on
the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed
by Monte Carlo sampling from the empirical metallicity distribution of
the Galactic halo from \cite{ivezic08} (centered around [Fe/H] $\sim$
-1.5). Within 2~kpc, Gaia will measure distances to these stars with
better than 2\% accuracy --- RR Lyrae in this volume will be used to
test and calibrate the Spitzer PL relation described above. Beyond the
2~kpc threshold, the mid-IR PL relation for RR Lyrae will provide more
accurate distance measurements.

@@

@@

Combined Spitzer and Gaia measurements of distances and proper motions
of RR Lyraes in the Sgr debris will open up new avenues of
interpretation. A 2\% distance error at $\sim$30~kpc
($\epsilon_{d}=0.6~{\rm kpc}$) is smaller than the predicted thickness
of the stream, ranging from $\sim$2-$10$~kpc. Combined with Gaia
tangential velocity measurement accuracies ($\epsilon_{v}\sim2$~km/s
at 30~kpc) less than the intrinsic dispersion for much of the stream
\citep[$\sigma_v\sim10$~km/s;][]{majewski04}, and ground-based radial
velocities, the measured phase-space position of each individual RR
Lyrae becomes a powerful probe of the potential. The next section
outlines a new method to take advantage of this information.

\section{Description and test of our algorithm}
\label{sec:method}
With access to full phase-space information for stars in a tidal
stream, it is possible to use the individual stars as potential
measures by exploting the fact that the stars must have come from the
same progenitor. If the present-day phase-space position of the
progenitor system is also known, this places an intuitive constraint
on the potential of the Milky Way: if the orbits of the stars and the
orbit of the progenitor are integrated \emph{backwards} in time
(imagine watching satellite destruction in ``rewind"), the stars
should recombine with the progenitor. If the potential is incorrect,
the orbits of the stars will diverge from that of the progenitor and
thus will not be recaptured by the satellite system
(Figure~\ref{fig:ps_distance}).

This approach was originally proposed by \citep{johnston99a} and was
designed and tested for the proposed characteristics of the Space
Interferometry Mission \citep{unwin08}, which was expected to be able
to measure proper motions with $\mu$as/year accuracy for stars as
faint as $V=20$. Distances to stream members would only be poorly
known and were estimated in the method by using anticipated properties
of tidal debris. Below we present an updated version of the algorithm:
the promise of 2\% distances to RR Lyrae stars (see Section
\ref{sec:spitzer}) enables a direct measurement (rather than
approximate estimate) of the position of a star within its debris
structure. The test statistic that quantifies how well stars recombine
with the satellite has been rigorously redefined.

@@

\subsection{The algorithm: Rewinder}
Quantifying this method requires a sample of stars with known full
space kinematics $(\bs{x}_{i}, \bs{v}_{i})|_{t=0}$ (e.g., measurements
of all position and velocity components for these stars \emph{today}
at $t=0$), the orbital parameters for the progenitor system
$(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a functional form for the
potential, $\Phi({\boldsymbol\theta})$. For a given set of potential
parameters, $\boldsymbol\theta$, the orbits of the stars and
progenitor are integrated backwards for several Gigayears (determined
from an estimate of the interaction time of the system). The stars and
progenitor system are treated as test particles, so that the debris
stars do not become physically re-bound to the satellite. Instead, at
each timestep $t_j$, for each particle $i$, a set of normalized,
relative phase-space coordinates are computed
\begin{equation}
  \bs{q}_{i} = \frac{\bs{x}_{i} -
    \bs{x}_{p}}{R_{\rm tide}}\,\,\,\,,\,\,\,\,\bs{p}_{i} = \frac{\bs{v}_{i} -
    \bs{v}_{p}}{v_{\rm esc}}
\end{equation}
where $(\bs{x},\bs{v})_{i}$ and $(\bs{x},\bs{v})_{p}$ are the
phase-space coordinates for the particles and progenitor,
respectively. These definitions require an estimate of the mass of the
satellite, $m_{sat}$, which, combined with the orbital radius of the
satellite, $R$, and the computed enclosed mass of the potential within
$R$, $M_{enc}$, sets the instantaneous tidal radius and escape
velocity,
\begin{equation}
  R_{\rm tide}=R\Big(\frac{m_{\rm sat}}{3M_{\rm enc}}\Big)^{1/3}\,\,\,\,,\,\,\,\,
  v_{\rm esc}=\sqrt{\frac{2Gm_{\rm sat}}{R_{\rm tide}}}.
\end{equation}
Qualitatively, when the distance in this normalized six-dimensional
space, $D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} <
\sqrt{2}$, the star is likely recaptured by the
satellite. \citet{johnston99a} (hereafter JZSH) imposed a similar
condition as a hard boundary and used the number of recaptured
particles in a given backwards-integration as a cost function for
optimization. What follows is a description of an updated procedure
with a more natural, statistically-motivated choice for an objective
function.

@@

For each star, $i$,
the phase-space distance, $D_{\rm ps}$, is computed at each timestep
$t_{j}$, and the vector with the minimum phase-space distance is stored
\begin{align}
  t^*_{i} &= \argmin_{t} D_{{\rm ps},i}\\
  \bs{A}_{i} &= (\bs{q}_{i}(t^*_{i}),\bs{p}_{i}(t^*_{i})).
\end{align}
Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space
distance vectors for each star, where $k\in[1,6]$. Intuitively, the
variance of the distribution of minimum phase-space vectors will be
larger for orbits integrated in an incorrect potential relative to the
distribution computed from the `true' orbital history of the stars: in
an incorrect potential, small differences in the orbits of the stars
relative to the orbit of the progenitor cause them to spread out in
this relative, normalized phase space. Thus, the \emph{generalized
  variance} of the distribution --- computed for a given set of
potential parameters, $\bs{\theta}$ --- is a natural choice for the
scalar objective function used in constraining the potential of the
Milky Way
\begin{align}
  \Sigma_n &= \mathrm{Cov}( \bs{A}_{ik}) \\
  f(\bs{\theta}) &= \ln \det \Sigma.
\end{align}

[Still working on this bit...need to figure out best way to express
  what I'm trying to say.] Though this is not explicitly a
\emph{likelihood}, optimizing over this statistic is a very simple
approximation to performing a maximum-likelihood
optimization. Assuming the samples in $\bs{A}_{ik}$ are approximately
normal-distributed, the variance is closely related to the Shannon
entropy \citep{shannon1948}, $H$,
\begin{equation}
  H \propto \log \sigma .
\end{equation}
Minimizing the entropy is then equivalent to minimizing the loss of
\emph{information} for a given choice of parameters, $\bs{\theta}$,
given the ``data,'' $\bs{A}_{ik}$, and is thus related to
maximum-likelihood estimation \citep[see][for more
  detail]{mackay2002}.

\subsection{Application to Simulated Data} \label{sec:results}
The LM10 simulation data (see Section \ref{sec:sgr}) is a perfect
test-bed for evaluating the effectiveness of this method. We start by
extracting both particle data and the satellite orbital parameters
from the present-day snapshot of the simulation
data.\footnote{\url{www.astro.virginia.edu/~srm4n/Sgr/data.html}} We
then `observe' a sample of 100 stars from the first leading and
trailing wraps of the stream: the radial velocity errors are drawn
from a Gaussian $\epsilon_{rv} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm
  km/s})$, the tangential velocity errors depend on the distance to
the star and are computed from the expected Gaia error curve (shown in
Figure~\ref{fig:gaia_errors}), and the distance errors on a star of
distance $D$ are $\epsilon_{dist} \sim \mathcal{N}(0,0.02\times
D)$. Figure~\ref{fig:variance_proj} shows two-dimensional projections
of the six-dimensional minimum phase-space distance distribution for
the observed particles in the correct and an incorrect potential: it
is clear that the variance of the distribution increases in the wrong
potential. The generalized variance defines a convex hypersurface over
which we optimize four of the six logarithmic potential parameters:
$v_{circ}$, $\phi$, $q_1$, and $q_z$. We fix $q_2$ and $R_c$, which
are degenerate with combinations of the other
parameters. Figure~\ref{fig:objective} shows one-dimensional slices of
the objective function produced by varying each of the potential
parameters by $\pm10\%$ around the true values and holding all others
fixed.

@@

In anticipation of extending the above method to include a true
likelihood function, we use a parallelized Markov Chain Monte Carlo
(MCMC) algorithm \citep{foremanmackey2013} to sample from our
objective function.\footnote{Though MCMC is typically not an efficient
  optimization tool, in this case objective function is both noisy and
  expensive to compute. The stochasticity and easy parallelization of
  the algorithm mean it outperforms other optimizers on this problem.}
We use the median value of the (burnt-in) sample distribution as a
point estimate for the potential parameters. To assess the uncertainty
in the derived halo parameters, we bootstrap resample our 100 stars
100 times and estimate the potential parameters with each
resampling. Figure~\ref{fig:bootstrap} shows the recovered parameters
for each bootstrapped sample and demonstrates the power of the
combined Spitzer and Gaia data: a moderately sized sample of RR Lyrae
alone places quite strong constraints on the shape and mass of the
Galaxy's dark matter halo.

@@

\section{Discussion and future prospects}
\label{sec:discussion}

Many orbit fitting methods for measuring the Galactic potential assume
that the stars in a given stream are on the same
orbit. \cite{sanders13a} show that this assumption can lead to nearly
100\% errors in the estimated potential parameters, even without
realistic observational errors on the kinematic data. The
\emph{Rewinder} is free of this assumption and instead \emph{relies}
on the fact that each star is on a different orbit, and is thus an
independent potentiometer. Qualitatively, the best potential is the
one that gets the most of these orbits to intersect the progenitor
over the history of interaction with the system.

The strengths of this method stem from its simplicity: it requires
only a rough estimate of the satellite mass $m_{\rm sat}$ combined
with backwards integration of orbits. There are no assumptions made
about, or N-body modeling of, the internal distribution of satellite
stars. Thus, the \emph{Rewinder} is applicable to any debris that is
known to come from a single object and not restricted to the coldest
tidal streams. In principle, the method could also be applied to the
vast debris {\it clouds} that have been discovered over large areas of
the sky \citep[e.g. the Triangulum Andromeda and Hercules-Aquila
  clouds]{rochapinto04,belokurov06}, or even stars that have only
associations in orbital properties and do not form a coherent spatial
structure \citep[such as the angular momentum groupings in local
  giants found by][]{helmi99}. The method trivially extends to
combining constraints from multiple debris systems at once by simply
integrating all debris from several satellites simultaneously, with
$D_{\rm ps}$ defined appropriately for each star.

It is also important to characterize the the limitations of this
method. Firstly, the measurement errors for RR Lyraes associated with
the very coldest streams \citep[e.g., the globular clusters Pal5 and
  GD1;][]{odenkirchen02,koposov10} will likely be too large to resolve
the minute differences in orbital properties between the debris and
satellite to be detected. Second, the current version of the algorithm
relies on accurate knowledge of the current position and velocity of
the parent satellite, which may not be available in some cases
\citep[for example, the Orphan Stream;][]{belokurov07}. Lastly, the
present prescription treats the progenitor as a point-mass during
orbit integration. Second-order effects in orbit evolution (e.g.,
dynamical friction) are currently ignored, leading to minor biases in
the estimated potential parameters (for example, XXX in
Figure~\ref{fig:bootstrap}).

[KVJ: Ok, there is a question as to whether we even want to say something like the below...?]

@@

\section{Conclusions}
\label{sec:conclusion}

This paper presents an algorithm for measuring the Galactic potential
that anticipates combined data from the Spitzer and Gaia satellite
missions that promise precise, full phase-space measurements of RR
Lyrae stars in the halo of our Galaxy. When applied to a sample of 100
stars (with realistic observational errors) sampled from the
\cite{law10} N-body simulation of the destruction of the Sgr dwarf
satellite, the method recovers the shape and orientation of the dark
matter potential XXX. This success provides strong motivation for 1)
further theoretical work to develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by the \emph{Rewinder}, 2) an investigation of the power of
using multiple, complimentary debris structures (e.g., with different
morphologies and orientations) in building a comprehensive picture of
the Galactic mass distribution, and 3) a Spitzer survey of RR Lyrae
stars in debris structures around the Milky Way to get precise
distances to combine with near-future Gaia velocity data. We conclude by noting that this method is complimentary to the orbit fitting algorithm presented in \citet{sanders13b} by favoring warmer, more massive debris structures like Sgr. By developing and comparing independent methods for measuring the potential of the Milky Way, we will begin to develop a deeper understanding of the vast structures of dark matter that permeate the universe. [Yea, the last sentence I wrote in the late night and am probably delirious...]

>>>>

\subsection{Gaia and the age of astrometry}
\label{sec:gaia}
The Gaia satellite \citep{gaia01}, planned for launch in late 2013, is
an astrometric mission which aims to measure the positions of billions
of stars with 10-100~$\mu$as accuracies. With over 50 observations of
the entire sky during the five-year lifetime of the mission, Gaia will
revolutionize the field of Galactic Astronomy by providing sufficient
proper motion accuracy to create full six-dimensional phase-space maps
of the Galaxy with $<$10\% distance errors for heliocentric distances of
up to $\sim$6~kpc for RR Lyrae stars.

Figure~\ref{fig:gaia_errors} shows the Gaia end-of-mission distance
and tangential velocity error estimates for RR Lyrae. The velocity
uncertainty depends on the magnitude (and color) of the source, which depends on
the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed
by Monte Carlo sampling from the empirical metallicity distribution of
the Galactic halo from \cite{ivezic08} (centered around [Fe/H] $\sim$
-1.5). Within 2~kpc, Gaia will measure distances to these stars with
better than 2\% accuracy --- RR Lyrae in this volume will be used to
test and calibrate the Spitzer PL relation described above. Beyond the
2~kpc threshold, the mid-IR PL relation for RR Lyrae will provide more
accurate distance measurements.

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/spitzer_proposal/lm10.png}
\caption{ todo }\label{fig:lm10}
\end{center}
\end{figure}

@@

Combined Spitzer and Gaia measurements of distances and proper motions
of RR Lyraes in the Sgr debris will open up new avenues of
interpretation. A 2\% distance error at $\sim$30~kpc
($\epsilon_{d}=0.6~{\rm kpc}$) is smaller than the predicted thickness
of the stream, ranging from $\sim$2-$10$~kpc. When combined with Gaia proper motion errors,
tangential velocity measurement accuracies ($\epsilon_{v}\sim2$~km/s
at 30~kpc; see Figure~\ref{fig:gaia_errors}) are less than the intrinsic dispersion for much of the stream
\citep[$\sigma_v\sim10$~km/s;][]{majewski04} and the measured phase-space position of each individual RR
Lyrae becomes a powerful probe of the potential. The next section
outlines a new method to take advantage of this information.

\section{Description and test of our algorithm}
\label{sec:method}
With access to full phase-space information for stars in a tidal
stream, it is possible to use the individual stars as potential
measures by exploting the fact that the stars must have come from the
same progenitor. If the present-day phase-space position of the
progenitor system is also known, this places an intuitive constraint
on the potential of the Milky Way: if the orbits of the stars and the
orbit of the progenitor are integrated \emph{backwards} in a a potential that accurately models the Milky Way
(imagine watching satellite destruction in ``rewind"), the stars
should recombine with the progenitor. If the potential is incorrect,
the orbits of the stars will diverge from that of the progenitor and
thus will not be recaptured by the satellite system
(Figure~\ref{fig:ps_distance}).

This approach was originally proposed by \citet{johnston99a} and was
designed and tested for the proposed characteristics of the Space
Interferometry Mission \citep{unwin08}, which was expected to be able
to measure proper motions with $\mu$as/year accuracy for stars as
faint as $V=20$. Distances to stream members would only be poorly
known and were estimated in the method by using anticipated properties
of tidal debris. Below we present an updated version of the algorithm:
the promise of 2\% distances to RR Lyrae stars (see Section
\ref{sec:spitzer}) enables a direct measurement (rather than
approximate estimate) of the position of a star within its debris
structure; the test statistic that quantifies how well stars recombine
with the satellite has been rigorously redefined.

@@

\subsection{The algorithm: Rewinder}
Quantifying this method requires a sample of stars with known full
space kinematics $(\bs{x}_{i}, \bs{v}_{i})|_{t=0}$ (e.g., measurements
of all position and velocity components for these stars \emph{today}
at $t=0$), the orbital parameters for the progenitor system
$(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a functional form for the
potential, $\Phi({\boldsymbol\theta})$. For a given set of potential
parameters, $\boldsymbol\theta$, the orbits of the stars and
progenitor are integrated backwards for several Gigayears (determined
from an estimate of the interaction time of the system). The stars and
progenitor system are treated as test particles, so that the debris
stars do not become physically re-bound to the satellite. Instead, at
each timestep $t_j$, for each particle $i$, a set of normalized,
relative phase-space coordinates are computed
\begin{equation}
  \bs{q}_{i} = \frac{\bs{x}_{i} -
    \bs{x}_{p}}{R_{\rm tide}}\,\,\,\,,\,\,\,\,\bs{p}_{i} = \frac{\bs{v}_{i} -
    \bs{v}_{p}}{v_{\rm esc}}
\end{equation}
where $(\bs{x},\bs{v})_{i}$ and $(\bs{x},\bs{v})_{p}$ are the
phase-space coordinates for the particles and progenitor,
respectively. These definitions require an estimate of the mass of the
satellite, $m_{sat}$, which, combined with the orbital radius of the
satellite, $R$, and the computed enclosed mass of the potential within
$R$, $M_{enc}$, sets the instantaneous tidal radius and escape
velocity,
\begin{equation}
  R_{\rm tide}=R\Big(\frac{m_{\rm sat}}{3M_{\rm enc}}\Big)^{1/3}\,\,\,\,,\,\,\,\,
  v_{\rm esc}=\sqrt{\frac{2Gm_{\rm sat}}{R_{\rm tide}}}.
\end{equation}
Qualitatively, when the distance in this normalized six-dimensional
space, $D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} <
\sqrt{2}$, the star is likely recaptured by the
satellite. \citet{johnston99a} imposed a similar
condition as a hard boundary and used the number of recaptured
particles in a given backwards-integration as a cost function for
optimization. What follows is a description of an updated procedure
with a more natural, statistically-motivated choice for an objective
function.

@@

For each star, $i$,
the phase-space distance, $D_{\rm ps}$, is computed at each timestep
$t_{j}$, and the vector with the minimum phase-space distance is stored
\begin{align}
  t^*_{i} &= \argmin_{t} D_{{\rm ps},i}\\
  \bs{A}_{i} &= (\bs{q}_{i}(t^*_{i}),\bs{p}_{i}(t^*_{i})).
\end{align}
Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space
distance vectors for each star, where $k\in[1,6]$. Intuitively, the
variance of the distribution of minimum phase-space vectors will be
larger for orbits integrated in an incorrect potential relative to the
distribution computed from the `true' orbital history of the stars: in
an incorrect potential, small differences in the orbits of the stars
relative to the orbit of the progenitor cause them to spread out in
this relative, normalized phase space. Thus, the \emph{generalized
  variance} of the distribution --- computed for a given set of
potential parameters, $\bs{\theta}$ --- is a natural choice for the
scalar objective function, $f(\bs{\theta})$, used in constraining the potential of the
Milky Way
\begin{align}
  \Sigma_n &= \mathrm{Cov}( \bs{A}_{ik}) \\
  f(\bs{\theta}) &= \ln \det \Sigma.
\end{align}

%[Still working on this bit...need to figure out best way to express
%  what I'm trying to say.] Though this is not explicitly a
%\emph{likelihood}, optimizing over this statistic is a very simple
%approximation to performing a maximum-likelihood
%optimization. Assuming the samples in $\bs{A}_{ik}$ are approximately
%normal-distributed, the variance is closely related to the Shannon
%entropy \citep{shannon1948}, $H$,
%\begin{equation}
%  H \propto \log \sigma .
%\end{equation}
%Minimizing the entropy is then equivalent to minimizing the loss of
%\emph{information} for a given choice of parameters, $\bs{\theta}$,
%given the ``data,'' $\bs{A}_{ik}$, and is thus related to
%maximum-likelihood estimation \citep[see][for more
%  detail]{mackay2002}.

\subsection{Application to Simulated Data} \label{sec:results}
The LM10 simulation data (see Section \ref{sec:sgr}) is a perfect
test-bed for evaluating the effectiveness of this method. We start by
extracting both particle data and the satellite orbital parameters
from the present-day snapshot of the simulation
data.\footnote{\url{www.astro.virginia.edu/~srm4n/Sgr/data.html}} We
then `observe' a sample of 100 stars from the first leading and
trailing wraps of the stream; the radial velocity and distance errors are drawn % APW HERE
from Gaussians ($\varepsilon_{\rm RV} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm
  km/s})$ and $\varepsilon_{\rm D} \sim \mathcal{N}(0,0.02\times
D)$) and the proper motion errors are computed from the expected Gaia error curve. Figure~\ref{fig:variance_proj} shows two-dimensional projections
of the six-dimensional minimum phase-space distance distribution for
the observed particles in the correct and an incorrect potential: it
is clear that the variance of the distribution increases in the wrong
potential. The generalized variance defines a convex hypersurface over
which we optimize four of the six logarithmic potential parameters:
$v_{circ}$, $\phi$, $q_1$, and $q_z$. We fix $q_2$ and $R_c$, which
are degenerate with combinations of the other
parameters. Figure~\ref{fig:objective} shows one-dimensional slices of
the objective function produced by varying each of the potential
parameters by $\pm10\%$ around the true values and holding all others
fixed.

@@

In anticipation of extending the above method to include a true
likelihood function, we use a parallelized Markov Chain Monte Carlo
(MCMC) algorithm \citep{foremanmackey2013} to sample from our
objective function.\footnote{Though MCMC is typically not an efficient
  optimization tool, in this case objective function is both noisy and
  expensive to compute. The stochasticity and easy parallelization of
  the algorithm mean it outperforms other optimizers on this problem.}
We use the median value of the (burnt-in) sample distribution as a
point estimate for the potential parameters. To assess the uncertainty
in the derived halo parameters, we bootstrap resample our 100 stars
100 times and estimate the potential parameters with each
resampling. Figure~\ref{fig:bootstrap} shows the recovered parameters
for each bootstrapped sample and demonstrates the power of the
combined Spitzer and Gaia data: a moderately sized sample of RR Lyrae
alone places quite strong constraints on the shape and mass of the
Galaxy's dark matter halo. [APW: explicitly state limits here, 1-sigma bounds]

\begin{figure}[h]
\begin{center}
\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/spitzer_proposal/bootstrap.pdf}
\caption{ todo }\label{fig:bootstrap}
\end{center}
\end{figure}

@@

\section{Discussion, strengths, and limitations}
\label{sec:discussion}

The strengths of this method stem from its simplicity: it requires
only a rough estimate of the satellite mass $m_{\rm sat}$ combined
with backwards integration of orbits. Many orbit fitting methods for measuring the Galactic potential assume
that the stars in a given stream are on the same
orbit. \cite{sanders13a} show that this assumption can lead to nearly
100\% errors in the estimated potential parameters, even without
realistic observational errors on the kinematic data. The
\emph{Rewinder} is free of this assumption and instead \emph{relies}
on the fact that each star is on a different orbit, and is thus an
independent potentiometer. There are also no assumptions made
about, or N-body modeling of, the internal distribution of satellite
stars. Thus, the \emph{Rewinder} is applicable to any debris that is
known to come from a single object and not restricted to the coldest
tidal streams. In principle, the method could also be applied to the
vast debris {\it clouds} that have been discovered over large areas of
the sky \citep[e.g. the Triangulum Andromeda and Hercules-Aquila
  clouds]{rochapinto04,belokurov06}, or even stars that have only
associations in orbital properties and do not form a coherent spatial
structure \citep[such as the angular momentum groupings in local
  giants found by][]{helmi99}. The method trivially extends to
combining constraints from multiple debris systems at once by simply
integrating all debris from several satellites simultaneously, with
$D_{\rm ps}$ defined appropriately for each star.

It is also important to characterize the the limitations of this
method. Firstly, the measurement errors for RR Lyraes associated with
the very coldest streams \citep[e.g., the globular clusters Pal5 and
  GD1;][]{odenkirchen02,koposov10} will likely be too large to resolve
the minute differences in orbital properties between the debris and
satellite to be detected. Second, the current version of the algorithm
relies on accurate knowledge of the current position and velocity of
the parent satellite, which may not be available in some cases
\citep[for example, the Orphan Stream;][]{belokurov07}. Lastly, the
present prescription neglects second-order effects in orbit evolution (e.g.,
dynamical friction) which may lead to biases in
the estimated potential parameters.

\section{Conclusions and motivation for future work}
\label{sec:conclusion}

This paper presents an algorithm for measuring the Galactic potential
that anticipates combined data from the Spitzer and Gaia satellite
missions that promise precise, full phase-space measurements of RR
Lyrae stars in the halo of our Galaxy. When applied to a sample of 100
stars (with realistic observational errors) sampled from the
\cite{law10} N-body simulation of the destruction of the Sgr dwarf
satellite, the method recovers the shape and orientation of the dark
matter potential XXX.

@@

This success provides strong motivation for 1)
further theoretical work to develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by the \emph{Rewinder}, 2) an investigation of the power of using multiple, complimentary debris structures (e.g., with different
morphologies and orientations) in building a comprehensive picture of
the Galactic mass distribution, and 3) a Spitzer survey of RR Lyrae
stars in debris structures around the Milky Way to get precise
distances to combine with near-future Gaia velocity data.
